---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, we will load the required libraries.  Second we must read and summarize the code, in order to draw the histogram of steps taken each day.
```{r}
library(stringr)
library(dplyr)
library(ggplot2)
raw <- read.csv("C:/Users/Pierre/Documents/Github/RepData_PeerAssessment1/activity/activity.csv")
by_date <- group_by(raw, date)
steps_by_day <- summarize(by_date, total_steps = sum(steps))
ggplot(steps_by_day, aes(total_steps)) + geom_histogram()
```


## What is mean total number of steps taken per day?
```{r}
daily_mean <- as.numeric(mean(steps_by_day$total_steps, na.rm = TRUE))
daily_median <- median(steps_by_day$total_steps, na.rm = TRUE)
```
The mean of the daily steps taken is `r daily_mean` and the median is `r daily_median`.

## What is the average daily activity pattern?
```{r}
by_interval <- group_by(raw,interval)
avg_steps_per_interval <- summarize(by_interval, steps = mean(steps, na.rm = TRUE))
ggplot(avg_steps_per_interval, aes(interval, steps)) + geom_line()
max_interval_avg <- avg_steps_per_interval[avg_steps_per_interval$steps == max(avg_steps_per_interval$steps),]
```
The interval containing the maximum, on average, activity is `r max_interval_avg$interval`, which has a value of `r max_interval_avg$steps`.

## Imputing missing values
```{r}
rows_with_nas <- sum(!complete.cases(raw))
missing_steps <- sum(is.na(raw$steps))
```
There are `r rows_with_nas` NAs in the data and `r missing_steps` of them occur in the steps column.   My strategy to replace missing values is to write a zero in for every missing value.

```{r}

```

## Are there differences in activity patterns between weekdays and weekends?
